Partridge: An effective system for the automatic classification of the types of academic papers

### 2013-07-22 doc/bcspaper/article.tex
Removed ugly formatting from paper

\vspace{-15pt}
\section{Introduction} \label{sec:1}
\vspace{-11pt}

@@

\vspace{-15pt}
\section{Methods}
\label{sec:2}
%how types were chosen,

\subsection*{Collection of scientific articles} 
\vspace{-15pt}
Partridge allows users to upload any paper which is free of copyright
restrictions.  For the purpose of this study we needed a large set of papers
that we could label with CoreSC to investigate how this information assists
classification into paper type.  Open Access (OA) journals provide free read
access to their papers, but many do not permit the user of articles for data
mining purposes. The Public Library of Science (PLoS) journals contain large
volumes of OA literature under a permissive license that allows data mining.
They also use the PubMed Central markup schema, which is compatible with
SAPIENTA, for papers published through their journal. The PLoS journals
advanced search offers approximately 50 types of paper through which to
restrict the search.  Many of these paper type categories contain too few
papers to be useful, others are too ad hoc (e.g.  {\it Message from PLoS}).
Indeed it is not clear how these paper types have been identified. We chose to
look at a range of types, some of which we expect to overlap or have an unclear
distinction.  These are namely: Essay, Correspondence, Case Study, Perspective,
Viewpoint, Opinion, Review, Research.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.75\textwidth]{figures/corescs.png}
\caption{The CoreSC content of a research paper (left) and a review paper (right) randomly selected from the corpus}
\vspace{-20pt}
\label{fig:coresc_pies} 
\end{figure}

\subsection*{Paper Processing and classification}
%Discuss in more detail how Partridge processes papers (PDFX, sentence splitting) 
\vspace{-15pt}

@@

\vspace{-15pt}
\section{Results and Discussion}
\label{sec:3}
\vspace{-15pt}

@@

\vspace{-10pt}
\caption{Per-class recall, precision and F-measure micro-averaged over 10-fold cross validation, reporting the results on the held-out validation segment. Top paper types in bold.}

@@

\vspace{-20pt}
\label{tab:recallPrecision}       % Give a unique label
\end{table}

@@

\vspace{-10pt}

@@

\begin{tabular}{lllllllll}
\hline\noalign{\smallskip}
      &  Case Study   &     Corresp.      &   Essay  & Opinion &     Perspective    &     Research    &     Review  & Viewpoint \\
\noalign{\smallskip}\svhline\noalign{\smallskip}
Case Study    &    26   &     2     &   11    &    2  &      5 & 26       & 25 &       10 \\
Correspondence  &      4  &      54   &     3    &    9     &   15 & 8   &     4 &       2 \\
Essay      &  8     &   5    &   90    &    2 &       39      &  5 & 46    &    5 \\
Opinion    &    4   &     7 &       2   &     23      &  27     &   10   &  3  &      17 \\
Perspective &       11  &      16    &    49       & 30      &  51 & 6      &  26    &   11 \\
Research      &  19    &    6 &       2 &       8    &    5      &  140  &   12    &    8\\
Review       & 27     &   8     &   35    &    5      &  13    &    21   &  197 &       6\\
Viewpoint    &    7  &      9   &     3     &   19     &   13 & 10     &   2  &      11 \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
\vspace{-15pt}
\end{table}

@@

\vspace{-15pt}
\section{Summary and Conclusions}
\label{sec:4}
\vspace{-15pt}
To summarise, we have demonstrated that paper type can largely be predicted
from the distribution of the automatically obtained CoreSC sentence labels in
the full text of a paper. We have described some of the particular CoreSC
features that determine a paper type (e.g. Correspondence characterised by very
little Background, Opinions characterised by low percentage of Experiment) and
discussed which of the paper types are not easily separable in this way. We
recommended for example merging the Opinion, Viewpoint and Perspective articles
into a single category, which increased overall classification performance. It
is also potentially useful to distinguish between two types of Case study.
Analysis of an example tree shows the decision making process to be complex,
but agrees with our general understanding of paper types. Partridge allows refinement of paper search using CoreSC and paper type, both of which
can be intelligently determined using machine learning methods.

@@

\vspace{-10pt}
\begin{acknowledgement} We thank the Leverhulme trust for the support to Dr
Liakata's Early Career Fellowship and also EMBL-EBI, Cambridge UK for the
facilities offered to Dr Liakata.  \end{acknowledgement}
\vspace{-2pt}

\vspace{-30pt}
\bibliographystyle{plain}
\bibliography{partridge}

>>>>

\section{Introduction} \label{sec:1}

@@

\section{Methods}
\label{sec:2}
%how types were chosen,

\subsection*{Collection of scientific articles} 
Partridge allows users to upload any paper which is free of copyright
restrictions.  For the purpose of this study we needed a large set of papers
that we could label with CoreSC to investigate how this information assists
classification into paper type.  Open Access (OA) journals provide free read
access to their papers, but many do not permit the user of articles for data
mining purposes. The Public Library of Science (PLoS) journals contain large
volumes of OA literature under a permissive license that allows data mining.
They also use the PubMed Central markup schema, which is compatible with
SAPIENTA, for papers published through their journal. The PLoS journals
advanced search offers approximately 50 types of paper through which to
restrict the search.  Many of these paper type categories contain too few
papers to be useful, others are too ad hoc (e.g.  {\it Message from PLoS}).
Indeed it is not clear how these paper types have been identified. We chose to
look at a range of types, some of which we expect to overlap or have an unclear
distinction.  These are namely: Essay, Correspondence, Case Study, Perspective,
Viewpoint, Opinion, Review, Research.

@@

\begin{figure}[t]
\centering
\includegraphics[width=0.75\textwidth]{figures/corescs.png}
\caption{The CoreSC content of a research paper (left) and a review paper (right) randomly selected from the corpus}
\label{fig:coresc_pies} 
\end{figure}

\subsection*{Paper Processing and classification}
%Discuss in more detail how Partridge processes papers (PDFX, sentence splitting)

@@

\section{Results and Discussion}
\label{sec:3}

@@

\caption{Per-class recall, precision and F-measure micro-averaged over 10-fold cross validation, reporting the results on the held-out validation segment. Top paper types in bold.}

@@

\label{tab:recallPrecision}       % Give a unique label
\end{table}

@@

@@

\begin{tabular}{lllllllll}
\hline\noalign{\smallskip}
      &  Case Study   &     Corresp.      &   Essay  & Opinion &     Perspective    &     Research    &     Review  & Viewpoint \\
\noalign{\smallskip}\svhline\noalign{\smallskip}
Case Study    &    26   &     2     &   11    &    2  &      5 & 26       & 25 &       10 \\
Correspondence  &      4  &      54   &     3    &    9     &   15 & 8   &     4 &       2 \\
Essay      &  8     &   5    &   90    &    2 &       39      &  5 & 46    &    5 \\
Opinion    &    4   &     7 &       2   &     23      &  27     &   10   &  3  &      17 \\
Perspective &       11  &      16    &    49       & 30      &  51 & 6      &  26    &   11 \\
Research      &  19    &    6 &       2 &       8    &    5      &  140  &   12    &    8\\
Review       & 27     &   8     &   35    &    5      &  13    &    21   &  197 &       6\\
Viewpoint    &    7  &      9   &     3     &   19     &   13 & 10     &   2  &      11 \\
\noalign{\smallskip}\hline\noalign{\smallskip}
\end{tabular}
\end{table}

@@

\section{Summary and Conclusions}
\label{sec:4}
To summarise, we have demonstrated that paper type can largely be predicted
from the distribution of the automatically obtained CoreSC sentence labels in
the full text of a paper. We have described some of the particular CoreSC
features that determine a paper type (e.g. Correspondence characterised by very
little Background, Opinions characterised by low percentage of Experiment) and
discussed which of the paper types are not easily separable in this way. We
recommended for example merging the Opinion, Viewpoint and Perspective articles
into a single category, which increased overall classification performance. It
is also potentially useful to distinguish between two types of Case study.
Analysis of an example tree shows the decision making process to be complex,
but agrees with our general understanding of paper types. Partridge allows refinement of paper search using CoreSC and paper type, both of which
can be intelligently determined using machine learning methods.

@@

\begin{acknowledgement} We thank the Leverhulme trust for the support to Dr
Liakata's Early Career Fellowship and also EMBL-EBI, Cambridge UK for the
facilities offered to Dr Liakata.  \end{acknowledgement}

\bibliographystyle{plain}
\bibliography{partridge}
